A survey on NoSQL and NewSQL datastores

### 2014-11-26 NoSQL_tabel/NoSQL_table.tex
LSM tree & Bloom filters

\multicolumn{1}{|l|}{\textbf{Indexing}}                  & primary \& secondary (single field, compound, multikey, geospatial, text, hashed){[}2{]}                                                                        & Y, primary \& secondary                                                                                                  & Y, primary \& secondary                                                            & Y, primary \& secondary                                                             & Y, primary \& secondary{[}3{]}      & Y, primary \& secondary                                                  \\ \midrule
\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer + Shard-keys to speed up distributed queries{[}4{]}                                                                                             & Bloom filter{[}5{]} + No JOIN{[}6{]}                                                                                     & Bloom filter{[}7{]}                                                                & memcached                                                                           &                                     &                                                                           \\ \midrule

@@

\multicolumn{1}{|l|}{\textbf{Concurrency control}} & Atomic single document operations, otherwise 2-phase commit{[}10{]} Concurrent reads, exclusive writes (lock on DB level) & CQL offers CAS, row-level atomicity & Single-row transactions (ACID), OCC with MVCC for wider scope operations & Application can implement optimistic (using CAS) or pessimistic concurrency control &                                     & ACID + data access serialized and executed in single-threaded environment

>>>>

\multicolumn{1}{|l|}{\textbf{Indexing}}                  & primary \& secondary on every attribute, B-tree                                                                  & Y, primary \& secondary, LSM-tree                                                                                           & Y, primary \& secondary, LSM-tree                                                     & Y, primary \& secondary, B-tree                                                      & Y, primary \& secondary{[}3{]}      & Y, primary \& secondary                                                  \\ \midrule
\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer + Shard-keys to speed up distributed queries                                                                                             & Bloom filter{[}5{]}                                                                                     & Bloom filter{[}7{]}                                                                & memcached                                                                           &                                     &                                                                           \\ \midrule

@@

\multicolumn{1}{|l|}{\textbf{Concurrency control}} & Atomic single document operations, otherwise 2-phase commit; Concurrent reads, exclusive writes (lock on DB level) & CQL offers CAS, row-level atomicity & Single-row transactions (ACID possible), OCC with MVCC for wider scope operations & Application can implement optimistic (using CAS) or pessimistic concurrency control &                                     & ACID + data access serialized and executed in single-threaded environment \\ \midrule
\multicolumn{1}{|l|}{\textbf{Open-source?}}              & Y                                                                                                                                                               & Y                                                                                                                        & Y                                                                                  & Y                                                                                   & Y                                   & Y                                                                        

### 2014-11-26 paper_dec/paper_dec.tex
LSM tree & Bloom filters

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. As in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of data generated increases rapidly. The intermediary data generated in the sequencing pipeline for a human genome, ranges in the order of ?? TB. Also, a single human genome takes ?? MB/GB to store when sequenced,%TODO source
 so as this scales to genomes of millions of people and other organisms, storage requirements become ever more demanding in terms of scalability, latency and concurrency.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefor, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}
\end{itemize}

%Getalletjes over DNA-sequencing
%NoSQL datamodellen, CAP, NewSQL fifolosie
%Tabel?

@@

Because of the enormous choice of NoSQL and NewSQL datastores, an exhaustive study wasn't possible. This survey reviews the most popular datastores in a few relevant categories, namely document stores, wide columnar stores and NewSQL stores. A selection was then made based on similar criteria as in \cite{grolinger2013data}, following the ranking of DB-Engine Ranking \cite{db_engine_rank} as an indicator of popularity. This ranking tries to measure popularity based on a few parameters, such as the number of mentions on Web sites, general interest according to Google Trends, frequency of technical discussions on the Web, number of job offers, and number of professional profiles in which the systems are mentioned. The result were the document stores MongoDB and CouchBase Server, wide columnar stores Cassandra and HBase and NewSQL database VoltDB. There is already an existing application of MongoDB in the sequencing pipeline, making it even more relevant. Lastly, NewSQL query engine Cloudera Impala was also taken into consideration because of explicit interested from researchers in the lab. %TODO eerste vermelding van lab?
These 6 systems were then compared on characteristics relevant to high performance computing, such as indexing mechanisms, client interfaces to the data, distribution strategy, concurrency control and consistency models.

\include{tabel}

@@

Apache Cassandra, originally developped at Facebook but later open-sourced, combines the data-model of Google's BigTable system with the architecture and distribution strategy of Amazon's DynamoDB. It is intended for flexible, highly-available storage of very large datasets, running on cheap commodity hardware and offering high write throughput while not sacrificing read efficiency \cite{lakshman2010cassandra}.
\\
Since its inception Cassandra has however diverged slightly from the BigTable data model. It now offers tables and composite columns -much like in a conventional schema, and comes with its own query language, CQL\cite{cassandra_then&now}. CQL resembles SQL in many ways, but comes with some restrictions. For instance, it doesn't feature the JOIN clause \cite{cassandra_cql}. It strongly encourages physically collocating data that will be queried together, and supports denormalization with features such as collection types.\\
In order to scale linearly in the number of nodes to very large datasets, Cassandra operates in a fully masterless fashion. In terms of the CAP-theorem, it focusses on availability and partition-tolerance, rather than immediate consistency (though the user has control over the level of consistency, as will be explained). High availability and partition tolerance are achieved through asynchronous replication of rows over several nodes in the cluster, using consistent hashing and virtual nodes to handle high churn and incremental addition of nodes \cite{decandia2007dynamo} \cite{lakshman2010cassandra} \cite{cassandra_then&now}. The amount of replicas can be chosen by the client. Furthermore, Cassandra provides cross-datacenter replication to cope with entire datacenter failures.
\\On reads and writes, the client can specify the desired quorum, that is the number of replicas that acknowledge the operation. Although Cassandra was built with eventual consistency in mind, strong consistency can be obtained by choosing the quorum larger than the number of replicas \cite{grolinger2013data}.\\
In terms of concurrency control, Cassandra offers atomicity for single-row operations and serializable \textit{lightweight transactions}, essentially a compare-and-set functionality for larger operations \cite{cassandra_lightweight_trans}.\\
Being an open-source project, Cassandra is freely available, but there is an enterprise version with extra features such as integration with Apache Spark and Apache Solr, for complex analytical and search tasks \cite{cassandra_solr} \cite{cassandra_spark}.\\\\
In conclusion, Cassandra offers flexible data-modeling with decent querying and indexing support through its CQL-interface and scales incrementally to vast datasets, thanks to its extensive replication and failure-handling features.

@@

Apache HBase is an open source datastore using Google BigTable's datamodel, but running on top of the Hadoop Distributed File System (HDFS) instead of the Google File System (GFS).
\\Since its launch, HBase has adopted several secondary indexing mechanisms. HBase doesn't offer an SQL-like advanced querying language, but because of its HDFS underpinnings it can function as both input and output for Map/Reduce jobs.
\\The strong points of HBase are its strong consistency, which is rare among NoSQL-stores, and its concurrency model: ACID-compliant single row transactions and optimistic multi-version concurrency control for wider scope operations\cite{hbase_acid}\cite{grolinger2013data}\cite{borthakur2011apache}.\\\\
In conclusion, HBase offers flexible data-modeling and is especially useful when there is already a HDFS dataset and when Map/Reduce compatibility is desired. It scales well to very large datasets and offers excellent concurrency control and strong consistency.

>>>>

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. As in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of data generated increases rapidly. The intermediary data generated in the sequencing pipeline for a human genome, ranges in the order of ?? TB. Also, a single human genome takes ?? MB/GB to store when sequenced,%TODO source
 so as this scales to genomes of millions of people and other organisms, storage requirements become ever more demanding in terms of scalability, latency and concurrency.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefor, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}.
\item Document stores: these encapsulate key-value pairs in JSON or JSON-like documents.
\end{itemize}

%NoSQL datamodellen, CAP, NewSQL fifolosie
%Tabel?

@@

Because of the enormous choice of NoSQL and NewSQL datastores, an exhaustive study wasn't possible. This survey reviews the most popular datastores in a few relevant categories, namely document stores, wide columnar stores and NewSQL stores. A selection was then made based on similar criteria as in \cite{grolinger2013data}, following the ranking of DB-Engine Ranking \cite{db_engine_rank} as an indicator of popularity. This ranking tries to measure popularity based on a few parameters, such as the number of mentions on Web sites, general interest according to Google Trends, frequency of technical discussions on the Web, number of job offers, and number of professional profiles in which the systems are mentioned. The result were the document stores MongoDB and CouchBase Server, wide columnar stores Cassandra and HBase and NewSQL database VoltDB. There is already an existing application of MongoDB in DNA sequencing pipelines, making it even more relevant \cite{elprep_mongo}. Lastly, NewSQL query engine Cloudera Impala was also taken into consideration because of explicit interested from researchers in the lab. %TODO eerste vermelding van lab?
These 6 systems were then compared on characteristics relevant to high performance computing, such as indexing mechanisms, client interfaces to the data, distribution strategy, concurrency control and consistency models.

@@

Apache Cassandra, originally developped at Facebook but later open-sourced, combines the data-model of Google's BigTable system with the architecture and distribution strategy of Amazon's DynamoDB. It is intended for flexible, highly-available storage of very large datasets, running on cheap commodity hardware and offering high write throughput while not sacrificing read efficiency \cite{lakshman2010cassandra}.
\\
Since its inception Cassandra has however diverged slightly from the BigTable data model. It now offers tables and composite columns -much like in a conventional schema, and comes with its own query language, CQL\cite{cassandra_then&now}. CQL resembles SQL in many ways, albeit with some restrictions. For instance, it doesn't feature the JOIN clause \cite{cassandra_cql}. It strongly encourages physically collocating data that will be queried together, and supports denormalization with features such as collection types.\\
Cassandra has primary and secondary indexing mechanisms and implements these - like BigTable \cite{chang2008bigtable}, using log-structured merge trees (or LSM trees). These allow for deferring updates and flushing them to disk once enough have accumulated, thus reducing disk I/O. This greatly benefits write throughput \cite{o1996log}\cite{sears2012blsm}\cite{lakshman2010cassandra}. Cassandra, like BigTable, also offers Bloom filters\cite{mullin1983second}. These are an efficient probabilistic mechanism to predict whether an item is in a set (in this case, whether a key is in a table) and can thus significantly reduce unnecessary table scans\cite{lakshman2010cassandra}.\\
In order to scale linearly in the number of nodes to very large datasets, Cassandra operates in a fully masterless fashion. In terms of the CAP-theorem, it focusses on availability and partition-tolerance, rather than immediate consistency (though the user has control over the level of consistency, as will be explained). High availability and partition tolerance are achieved through asynchronous replication of rows over several nodes in the cluster, using consistent hashing and virtual nodes to handle high churn and incremental addition of nodes \cite{decandia2007dynamo} \cite{lakshman2010cassandra} \cite{cassandra_then&now}. The amount of replicas can be chosen by the client. Furthermore, Cassandra provides cross-datacenter replication to cope with entire datacenter failures.
\\On reads and writes, the client can specify the desired quorum, that is the number of replicas that acknowledge the operation. Although Cassandra was built with eventual consistency in mind, strong consistency can be obtained by choosing the quorum larger than the number of replicas \cite{grolinger2013data}.\\
In terms of concurrency control, Cassandra offers atomicity for single-row operations and serializable \textit{lightweight transactions}, essentially a compare-and-set functionality for larger operations \cite{cassandra_lightweight_trans}.\\
Being an open-source project, Cassandra is freely available, but there is an enterprise version with extra features such as integration with Apache Spark and Apache Solr, for complex analytical and search tasks \cite{cassandra_solr} \cite{cassandra_spark}.\\\\
In conclusion, Cassandra offers flexible data-modeling with decent querying and indexing support through its CQL-interface and scales incrementally to vast datasets, thanks to its extensive replication and failure-handling features.

@@

Apache HBase is an open source datastore using Google BigTable's datamodel, but running on top of the Hadoop Distributed File System (HDFS) instead of the Google File System (GFS).
\\Since its launch, HBase has adopted several secondary indexing mechanisms. For indexing, HBase also uses LSM trees\cite{borthakur2011apache}\cite{sears2012blsm}, and it also offers Bloom filters\cite{hbase_schema}. HBase doesn't offer an SQL-like advanced querying language, but because of its HDFS underpinnings it can function as both input and output for Map/Reduce jobs.
\\The strong points of HBase are its strong consistency, which is rare among NoSQL-stores, and its concurrency model: ACID-compliant single row transactions and optimistic multi-version concurrency control for wider scope operations\cite{hbase_acid}\cite{grolinger2013data}\cite{borthakur2011apache}.\\\\
In conclusion, HBase offers flexible data-modeling and is especially useful when there is already a HDFS dataset and when Map/Reduce compatibility is desired. It scales well to very large datasets and offers excellent concurrency control and strong consistency.

### 2014-11-26 paper_dec/tabel.tex
LSM tree & Bloom filters

\multicolumn{1}{|l|}{\textbf{Indexing}}                  & primary \& secondary on every attribute                                                                        & Y, primary \& secondary                                                                                                  & Y, primary \& secondary                                                            & Y, primary \& secondary                                                             & Y, primary \& secondary{[}3{]}      & Y, primary \& secondary                                                  \\ \midrule

>>>>

\multicolumn{1}{|l|}{\textbf{Indexing}}                  & primary \& secondary on every attribute, B-tree                                                                  & Y, primary \& secondary, LSM-tree                                                                                           & Y, primary \& secondary, LSM-tree                                                     & Y, primary \& secondary, B-tree                                                      & Y, primary \& secondary{[}3{]}      & Y, primary \& secondary                                                  \\ \midrule
